---
title: "a4_task2_nikole_vannest"
author: "Nikole Vannest"
date: "11/17/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(effsize)
library(broom)
library(palmerpenguins)
library(ggbeeswarm)
library(kableExtra)
library(lubridate)
library(scales
        )

hares <- read.csv(here("data", "bonanza_hares.csv"))
```

## Introduction:

## Data and analyses

## Exploratory findings

### Observations and descriptive statistics

```{r juvenile_cleanup, include = FALSE}
juvenile <- hares %>% 
  filter(age == "j") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) 

juvenile_table <- juvenile %>% 
  count(year) %>% 
  summarize(
    mean_count = mean(n),
    median_count = median(n),
    sd_count = sd(n),
    max_count = max(n),
    min_count = min(n)
  )
```


**Table 1.** Descriptive statistics of the total observations per year of juvenile snowshoe hares. 

```{r juvenile_summary, fig.align='center'}
juvenile_table %>% 
  kable(col.names = c("Mean count per year", "Median count per year", "Standard deviation", "Maximum per year", "Minimum per year")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

The distribution of observations per year of juvenile hares is fairly extreme. The standard deviation very large considering the maximum and minimum observations. There seems to be a fairly steep decline in number of observations of juvenile snowshoe hares from 1999 to 2012.

```{r basic_juvenile_plot, fig.align = "center"}
ggplot(juvenile, aes(x = year)) +
  geom_histogram(fill = "darkolivegreen4") +
  labs(x = "Year",
       y = "Juvenile Hare Count",
       title = "Number of Juvenile Hares Observed per year from 1999-2012") +
  scale_x_continuous(breaks = pretty_breaks(n = 12)) +
  scale_y_continuous(breaks = pretty_breaks(n = 7)) +
  theme_minimal()
```

**Figure 1.** Visual representation of the total number of observed juvenile hares from 1999 to 2012. The y axis depicts juvenile snowshoe hare count, with each year represented on the x axis. 

**Major Takeaways** The number of juvenile snowshoe hares observed per year across all species and sites are included to get a better understand of how the distribution of observations varies from year to year. The maximum number of observations = `r juvenile_table$max_count` was found early on in 1999, and the minimum in 2008 at `r juvenile_table$min_count` with no observations reported in 2002 and 2009. The mean count per year is `r juvenile_table$mean_count`, and median count is `r juvenile_table$median_count` with a standard deviation of `r juvenile_table$sd_count`. It could be interpreted from the figure above that the decline in the number of juvenile snowshoe hare observations per year from 1999 to 2012 could be due to a decline in adult snowshoe hare reproduction, or decreased population levels. 


  
